```{r head-0101, echo=FALSE}
h1("Conventions")
```

Just a few notes about the present tutorial. Most text will be formatted
as present, but at times there will be:
  
- linux command lines:
  
    `command-line-tool -option value -option value argument argument`

- R code without output:
  
```{r}
a <- 1
```

-   or with:
```{r}
(a <- 1)
```

-----

Finally, as this is a tutorial, you will be subjected to exercises. You should 
only proceed after answering the question asked. To answer the question, go to
_http://socrative.com_, select "Student Login" and use the room number you were
provided with. You will be taken directly to the active quizz, which you can 
answer at your own pace, however, you can answer a question only once. Once you
have submitted your answer, you can press "Close" in the present tutorial; the
current section of the tutorial will collapse (you can always re-open it, _e.g._
if you need clarifications about the question) and the next section will open. 
If anything were unclear at any time, do not hesitate to ask us (also contact us 
on slack).

