```{r head-0201, echo=FALSE}
h1("Introduction")
```
```{r head-0202, echo=FALSE}
h2("About the tutorial")
```

This tutorial adresses several aspect of an RNA-Seq data analysis. First, 
it deals with data pre-processing (a process rather agnostic to the type 
of NGS analysis, _i.e._ it would also be valid for ChIP-Seq or DNA-Seq).
That part of the tutorial introduces the use of the command line
(specifically the bourne-again shell: bash) and several software tools
to prepare sequencing data for downstrem analysis. Specifically, we will
be covering quality control, rRNA sorting, quality trimming, and expression 
quantification. For the latter, either by read mapping + feature summarization 
or by pseudo-alignment; or both . If these terms don’t mean much yet,
hopefully they will by the end of the practical. "Read mapping and feature
summarisation" deals with the manipulation of alignment files (in BAM format) 
and the obtention and manipulation of genomic (and genic) annotations to 
subsequently combine them to obtain a _raw count table_[^1]. The pseudo-alignment
method directly generates such a count table. This count table is the necessary 
minimal input for a Differential Expression analysis, which is the ultimate
step in this tutorial.

The goals of the tutorial are:

* First part
    + become familiar with the Linux command line
    + be able to carry out quality control and necessary pre-processing steps of sequencing data
* Second part
    + become familiar with the R environment
    + develop familiarity with R/Bioconductor packages for high-throughput analysis
* Third part
    + become familiar with the Differential Expression type of analysis

```{r head-0203, echo=FALSE}
  h2("Main softwares of interest")
```
1.  `FastQC` - Sequence data quality control

2.  `SortMeRNA` - rRNA filtering

3.  `Trimmomatic` - Low quality read and adapter trimming

4.  `STAR` - Read mapping

5.  `HTSeq` and `kallisto` - Feature summarization

```{r head-0204, echo=FALSE}
h2("Main R/Bioconductor packages of interest")
```

Bioconductor is a collection of R packages for the analysis and
comprehension of high-throughput genomic data. Among these, we will
focus on a few of them principally: `GenomicAlignments`,
`GenomicFeatures` and `GenomicRanges`. We will look at these packages (i) as an
example to understand what the process of __feature summarization__ is (_i.e._
taking a look under the hood) and (ii) as an introduction to computational optimisation in R. 

We will, in addition, use two packages for Differential Expression data 
analysis: `tximport ` and `DESeq2`.

```{r head-0205, echo=FALSE}
h3("A word on High-throughput sequence analysis")
```
  
Recent technological developments introduced high-throughput sequencing
approaches. A variety of experimental protocols and analysis workflows
address gene expression, regulation, and encoding of genetic variants.
Experimental protocols produce a large number (tens to hundreds of millions per
sample) of short (_e.g._: 35-250, single or paired-end) nucleotide sequences.
These are aligned to a reference or other genome or transcriptome.
Analysis workflows use the alignments to infer levels of gene expression
(RNA-seq), binding of regulatory elements to genomic locations
(ChIP-seq), or prevalence of structural variants (_e.g._: SNPs, short indels,
large-scale genomic rearrangements). Sample sizes range from minimal
replication (_e.g._: 2 samples per treatment group) to thousands of
individuals.

```{r head-0206, echo=FALSE}
    h3("A word on RNA-Seq")
```
RNA-Seq (RNA-Sequencing) has become the preferred method for measuring
gene expression, providing an accurate proxy for quantitation
of messenger RNA (mRNA) levels within a sample
[@Mortazavi:2008p740]. RNA-Seq has reached rapid maturity in data
handling, QC (Quality Control) and downstream statistical analysis
methods, taking substantial benefit from the extensive body of
literature developed on the analysis of microarray technologies and
their application to measuring gene expression. Although analysis of
RNA-Seq remains more challenging than for microarray data, the field has
now advanced to the point where it is possible to define mature
pipelines and guidelines for such analyses. However, with the exception
of commercial software options such as the CLCbio CLC Genomics
Workbench, for example, we are not aware of any fully integrated
open-source pipelines for performing these pre-processing steps. Both
the technology behind RNA-Seq and the associated analysis methods
continue to evolve at a rapid pace, and not all the properties of the
data are yet fully understood. Hence, the steps and available software
tools that could be used in such a pipeline have changed rapidly in
recent years and it is only recently that it has become possible to
propose a de-facto standard pipeline. Although proposing such a skeleton
pipeline is now feasible there remain a number of caveats to be kept in
mind in order to produce biologically and statistically sound results.

A pipeline, which we have developed for non-model plant organisms but is 
generally applicable can be seen in the figure below and a detailed 
description of it has been published[^2].

![The pre-processing pipeline used during the course][pipeline]

```{r head-0207, echo=FALSE}
    h2("A word on the dataset that is used in this tutorial")
```

As a running example, we use a dataset derived from a study performed in
_P. tremula_ [@Robinson:2014p6362]. The authors test
the evolutionary theory suggesting that males and females may evolve
sexually dimorphic phenotypic and biochemical traits concordant with
each sex having different optimal strategies of resource investment to
maximise reproductive success and fitness. Such sexual dimorphism should
result in sex biased gene expression patterns in non-floral organs for
autosomal genes associated with the control and development of such
phenotypic traits. Hence, the authors, among other approaches have
looked for gene expression differences between sex. This was achieved by
an RNA-Seq differential expression analysis between samples grouped by
sex. The samples have been sequenced on an Illumina HiSeq 2000, using
TruSeq adapters, through a 101 cycles paired-end protocol, which yielded
between 11 million and 34 million reads per sample.

In the following sections, we look at a subset of the data,
corresponding to reads obtained from individuals of the RNA-seq
experiment, and quantified against the _P. trichocarpa_ reference. 
The reason we used the  _P. trichocarpa_ reference is that we are currently 
establishing the _P. tremula_ genome. _P. trichocarpa_ is a closely related 
species (the latter present in north-western america
while the former grows in northern and central europe).

As a side note, reads were retrieved from the European Nucleotide
Archive (ENA, accession ID ERP002471).

```{r head-0208, echo=FALSE}
    h2("A word on Integrated Development Environment (IDE)")
```

There are numerous tools to support developing programs and softwares in _R_. 
For this course, we have selected one of them: the
[RStudio](http://rstudio.org) environment, which provides a
feature-full, user-friendly, cross-platform environment for working with _R_.
